---
title: "Automating our work away"
subtitle: "One consulting firm's experience with RMarkdown"
author: "Finbarr Timbers"
output:
    ioslides_presentation:
            css: assets/styles.css
            widescreen: yes
logo: assets/darkhorse_logo.png
---

<!--
As consultants, many of the projects that we work on are similar, with many
steps repeated verbatim across projects. Previously, our workflow was based
largely in Microsoft Office, with our analysis done manually in Excel, our
reports written in Word, and our presentations in Powerpoint. In 2015, we began
using R for much of our analysis, including making slide decks and reports in
RMarkdown. Our presentation discusses why we made the change, how we managed it,
and advice for other consulting firms looking to do the same.
-->

```{r, include=FALSE}
library(png)
library(grid)
library(ggplot2)
library(ggthemes)
library(scales)

knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(comment = FALSE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(fig.align = "center")

darkhorse_theme = function() {
  theme_bw() + theme(plot.background = element_rect(fill="transparent",
colour = NA),
panel.background = element_rect(fill = "transparent", colour = NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
panel.border = element_blank(),
legend.position="none",
axis.line.x = element_line(color = "black"),
axis.line.y = element_line(color = "black"),
axis.title.x=element_text(vjust=0.1),
plot.margin=unit(c(12.5, 1, 12.5, 1), "mm"))
#text=element_text(family="Open Sans"))
}
```

## Outline
### 1. About Darkhorse
### 2. Problems we faced
### 3. How we solved those problems
### 4. How you can improve R(Markdown) adoption

---

```{r fig.width=7.5, fig.height=5}
grid.raster(readPNG("assets/darkhorse_logo.png"))
```

---

```{r fig.width=7.5, fig.height=5}
grid.raster(readPNG("assets/histogram_1.png"))
```

---

```{r fig.width=7.5, fig.height=5}
grid.raster(readPNG("assets/histogram_2.png"))
```

---

```{r fig.width=7.5, fig.height=5}
grid.raster(readPNG("assets/histogram_3.png"))
```

---

```{r fig.width=7.5, fig.height=5}
grid.raster(readPNG("assets/histogram_4.png"))
```

---

```{r fig.width=7.5, fig.height=5}
grid.raster(readPNG("assets/line_chart_1.png"))
```

---

```{r fig.width=7.5, fig.height=5}
grid.raster(readPNG("assets/line_chart_2.png"))
```

---

```{r fig.width=7.5, fig.height=5}
grid.raster(readPNG("assets/line_chart_3.png"))
```

## Problems

## Problems
1. All styled differently. We need to edit every plot individually.

## Problems
1. All styled differently. We need to edit every plot individually.
2. No documentation on how to replicate.

## Problems
1. All styled differently. We need to edit every plot individually.
2. No documentation on how to replicate.
3. It takes time to replicate.

# Instead? R!

## R:
```
function darkhorse_histogram(x_data, x_var, x_breaks, title, 
                             y_breaks, col="#E69F00") {
    ggplot(x_data, aes(x_var)) +
        geom_bar(fill=col, aes(y=(..count..)/sum(..count..))) +
        geom_hline(yintercept= 0, colour=col) +
        darkhorse_theme() +
        ggtitle(title) +
        theme(axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks = element_blank()) +
        scale_y_continuous(breaks=y_breaks, labels=percent) +
        ylab("") + 
        geom_hline(yintercept=y_breaks[2:length(y_breaks)], color="white") +
        scale_x_continuous(breaks=x_breaks)
}
```

## Best practice charts

```{r}
darkhorse_histogram = function(x_data, x_var, x_breaks, title, 
                             y_breaks, col="#E69F00") {
    ggplot(x_data, aes_string(x_var)) +
        geom_bar(fill=col, aes(y=(..count..)/sum(..count..))) +
        geom_hline(yintercept= 0, colour=col) +
        darkhorse_theme() +
        ggtitle(title) +
        theme(axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            axis.ticks = element_blank()) +
        ylab("") + 
        scale_y_continuous(breaks=y_breaks, labels=percent) +
        geom_hline(yintercept=y_breaks[2:length(y_breaks)], color="white") +
        scale_x_continuous(breaks=x_breaks)
}

darkhorse_highlighted_linechart = function(line_data, x_var, y_var, 
                                           y_label, title, x_breaks, x_labels) { 
    ggplot(line_data, aes_string(x = x_var, y = y_var)) +
        geom_line(aes(colour = Colour, group = Group)) + 
        darkhorse_theme() +
        ylab(y_label) + ggtitle(title) +
        theme(axis.line.x = element_blank(),
              axis.line.y = element_blank(),
              axis.ticks = element_blank()) +
        scale_x_continuous(breaks=x_breaks, labels=x_labels)
}

y_breaks=seq(0, 0.30, by=0.05)
test_data = data.frame(1:100)
test_data$Region = c(rep.int(1, 25), 
                     rep.int(2, 15), 
                     rep.int(3, 10), 
                     rep.int(4, 25), 
                     rep.int(5, 20), 
                     rep.int(6, 5))
```
```{r fig.height=5, fig.width=9}
darkhorse_histogram(test_data, "Region", 
    x_breaks=1:6, 
    title="Regional distribution", 
    y_breaks=y_breaks, 
    col="#E69F00") 
```

## Best practice charts

```{r, fig.height=5, fig.width=9}
line_data = data.frame(
    Sales =  c(1100 + c(0, cumsum(runif(49, -20, 20))),
               1150 + c(0, cumsum(runif(49, -10, 10))),
               1200 + c(0, cumsum(runif(49, -30, 10))),
               1250 + c(0, cumsum(runif(49, -40, 10)))),
    Group = c(rep("1", 50), 
              rep("2", 50), 
              rep("3", 50), 
              rep("4", 50)),
    Month = rep(1:50, 4), 
    Colour = rep("grey", 200), stringsAsFactors=FALSE
)
line_data[line_data$Group == "3",]$Colour = "red"
months = seq(1, 360, 30)
intermediary_month_labels = as.Date(months, origin="2015/01/01", by="months", 
                                    format="%M")
month_labels = format(intermediary_month_labels, format="%b")
darkhorse_highlighted_linechart(line_data = line_data, 
    x_var = "Month", 
    y_var = "Sales", 
    y_label = "Sales ($)", 
    title = "", 
    x_breaks = seq(1, 48, 4), 
    x_labels = month_labels)
```

## What we did:

## What we did:
1. Use RMarkdown to write template ("best practice") slides.

## What we did:
1. Use RMarkdown to write template ("best practice") slides.
2. Make the slides accessible to the company (we used Git + Bitbucket).

## What we did:
1. Use RMarkdown to write template ("best practice") slides.
2. Made the slides accessible to the company (we used Git + Bitbucket).
3. Made it easy to run in other environments (we used installation scripts &
   Docker)

## What we did:
1. Use RMarkdown to write template ("best practice") slides.
2. Made the slides accessible to the company (we used Git + Bitbucket).
3. Made it easy to run in other environments (we used installation scripts &
   Docker)
4. Iterate (we used Git).

## What we did:
1. Use RMarkdown to write template ("best practice") slides.
2. Made the slides accessible to the company (we used Git + Bitbucket; 
we plan to publish a package soon).
3. Made it easy to run in other environments (we used installation scripts &
   Docker)
4. Iterate (we used Git).

## Lessons
1. Make it good.

## Lessons
1. Make it good.
2. Make it easy.

## Lessons
1. Make it good.
2. Make it easy.
3. Make it profitable.
